The illusory promise of the Aligned Rank Transform

### 2025-07-07 sandbox.qmd
minor correction

>>>>

**When the interpretation of effects becomes ambiguous.** Issues with interpreting interaction effects arise only when there are parallel main effects --- that is, when the coefficients of all interacting factors are non-zero. Interpretation issues may also extend to main effects when the interaction coefficient $a_{12}$ is non-zero. In such cases, we may observe a change in the marginal effect of a factor (e.g., $x_1$), even if its associated model coefficient (e.g., $a_1$) is zero. We discuss these scenarios in more depth in the [appendix](appendix.html#removable-main).

### 2025-07-07 text/10-intro.qmd
minor correction

However, given their substantial impact on research outcomes in our community, these widely adopted methods deserve closer scrutiny. As the first author previously argued [@tsandilas:2018], the HCI community often lacks the necessary expertise to critically assess the validity of statistical techniques. Our findings suggest that the widespread use of ART largely stems from a lack of awareness regarding its underlying assumptions. Many researchers treat ART as a general-purpose nonparametric alternative when ANOVA's assumptions are violated, overlooking the fact that ART’s alignment mechanism imposes even stricter assumptions --- including a linear relationship between the response and predictors, equal variances, and continuous data.

>>>>

However, given their substantial impact on research outcomes in our community, these widely adopted methods deserve closer scrutiny. As the first author previously argued [@tsandilas:2018], the HCI community lacks the necessary expertise to critically assess the validity of statistical techniques. Our findings suggest that the widespread use of ART largely stems from a lack of awareness of its underlying assumptions. Many researchers treat ART as a general-purpose nonparametric alternative when ANOVA's assumptions are violated, overlooking the fact that ART’s alignment mechanism imposes even stricter assumptions --- including a linear relationship between the response and predictors, equal variances, and continuous data.
